---
title: "Kish_HP2"
author: "John Kish"
format:
  html:
    embed-resources: true
---

```{r setup}

library(tidyverse)
library(here)

view(df)
```

employees of Montgomery County, MD paid in calendar year 2023

## Question 1 Gender



## Question 2 Total Compensation and Overtime



## Question 3 Department and Salary



## Question 4 Grade (w/ Focus on Management) and Salary





Some example questions could include "which gender has more overtime?", "how much of their total compensation is that overtime?", and "does this overtime and corresponding percentage of total compensation vary across department or decrease across time (here we only have access to "longevity pay" and not specific amount of years worked).










```{r codebook}

# Department
# Abbreivation/accronym for the department name. Example: ABS for the Alcohol Beverage Services department

# Department_Name
# the department name for all active, permanent employees of Montgomery County, MD (Maryland)


# Division
# The division for each department. That is, a single department can have multiple divisions and focuses.
# Example: ABS has the divisions that include those of ABS 85 Administrative Services
# and ABS 85 Beer Delivery Operations

# Gender
# the Gender of M and F. Assuming that since this is filings with the government, M is to be the legal sex of the individual M for Male and F for Female.

# Base_Salary
# precise base salaries (including down to the cent) for the individual, at the end of the calendar year.
# this is projected data. Should be accurate, but projection allows for greater flexibiliity.


# Overtime_Pay
# amount of overtime pay an employee/individual received for working beyond what they were required to do


# Longevity_Pay
# Longevity pay gives additional compensation for qualifying employees who have worked for a good deal of time. The longer time that is worked, the greater the longevity pay. Specifics about how long are dealt with in-house.


# Grade
# the standing an employee has within a company. The internal codes for all positions are not provided, but we can safely look at M1, M2, M3, etc. as those are clearly levels of management.

# M4-Senior Manager
# M3-Middle Manager
# M2-Middle Manager
# M1-First Level Manager


```





 You should ask/answer at least 4 significant questions in your analysis.
 
 By significant, I mean something meaningful about the data (not just randomly counting rows for no reason), and something that takes more than a couple of lines of code. 
 
 
 
 clear explanation of the questions you are exploring in introductory text
 
 
 
clear description of the datasets, similar to a data dictionary/codebook, but you don't need to describe variables you aren't using at all 
 
 
effective use of summaries, tables, and visualizations to address your questions, with explanations that refer to the key aspects of them
 
 
 
 not required, but very effective to use inline R code to refer to variables in the markdown text
 
 
working with files and projects
importing data

data cleaning and tidying rename variables, check for and resolve missing/malformed data as appropriate, set variable types

general data wrangling filtering/subsetting data, pivoting, joins - as needed

mutating, grouping, and summarizing

working with non-rectangular text data

explanation of variables inclusion of codebook/data dictionary and clear explanation of variables used in analysis, in the context of the chosen dataset

writing a quality data analysis report including text explanation of analysis, clear statement of questions and how answers are reported/visualized

targetted plots (at least 4) well-explained and contextualized in report, with appropriate labels and colors

sing GitHub and pushing changes


```{r import}


here::i_am("analysis/Kish_HP2.qmd")
library(here)

df <- read.csv(here("data/Employee_Salaries.csv"))


```

```{r}

```

