---
title: "Kish_HP2"
author: "John Kish"
format:
  html:
    embed-resources: true
---

```{r setup}

library(tidyverse)
library(here)


```


## Question 1



## Question 2



## Question 3



## Question 4





















 You should ask/answer at least 4 significant questions in your analysis.
 
 By significant, I mean something meaningful about the data (not just randomly counting rows for no reason), and something that takes more than a couple of lines of code. 
 
 
 
 clear explanation of the questions you are exploring in introductory text
 
 
 
clear description of the datasets, similar to a data dictionary/codebook, but you don't need to describe variables you aren't using at all 
 
 
effective use of summaries, tables, and visualizations to address your questions, with explanations that refer to the key aspects of them
 
 
 
 not required, but very effective to use inline R code to refer to variables in the markdown text
 
 
working with files and projects
importing data

data cleaning and tidying rename variables, check for and resolve missing/malformed data as appropriate, set variable types

general data wrangling filtering/subsetting data, pivoting, joins - as needed

mutating, grouping, and summarizing

working with non-rectangular text data

explanation of variables inclusion of codebook/data dictionary and clear explanation of variables used in analysis, in the context of the chosen dataset

writing a quality data analysis report including text explanation of analysis, clear statement of questions and how answers are reported/visualized

targetted plots (at least 4) well-explained and contextualized in report, with appropriate labels and colors

sing GitHub and pushing changes


```{r}


#previous example code from other hw - in progress
# write_rds(top_dogs_urls, "top_dogs_urls.rds")
# top_dogs_urls <- read_rds("top_dogs_urls.rds")

here::i_am("hp2-kishjohn/analysis")
library(here)

# write_csv(data2, "Employee_Salaries.csv")
data <- read.csv(here("data/Employee_Salaries.csv")
                 
# data_removed_part <- filter(data, Hash != '9dAvrH0+R6a0U5adPzZSyA')

getwd()

```

```{r}

```

